\pdftitle

### 2017-05-18 maxNt.tex
discussion

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-18T13:36:45+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

 \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}

@@

% \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}

@@

\newcommand*{\yxx}{x}%subpop state
\newcommand*{\yx}{\bm{\yxx}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yX}{\bm{X}}%pop state
\newcommand*{\yy}{\bm{y}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
\newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
\newcommand*{\yr}{\bm{r}}%subpop value
\newcommand*{\ys}{\bm{s}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{R}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varIota}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varIota'}
\newcommand*{\yHb}{\varIota''}
\newcommand*{\yHc}{\varKappa}
\newcommand*{\yHd}{\Hat{\varIota}}
%Lagr multipliers
\newcommand*{\yf}{\bm{f}}
\newcommand*{\yc}{\bm{c}}
\newcommand*{\yL}{\bm{\varLambda}}
\newcommand*{\yl}{\bm{\lambda}}
\newcommand*{\yk}{\kappa}
\newcommand*{\yK}{\varKappa}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

@@

A perfunctory intuitive reasoning seems insufficient for clarifying this
point. Suppose we do not know whether the sample is really part of a larger
population; we do not know if $N=n$ or how large is $N$ otherwise. Call
this state of ignorance $\yHd$. From the
point of view of probability theory, th
 ignorance about the
value of $N$ is expressed by assigning a probability distribution
$\p(N \| \yH)$ that vanishes if $N<n$, since we know that $N\ge n$. Let's
call thi
Probability assignments for the sample state that do not assume a specific
value of $N$ are then obtained by marginalization

The former application must therefore implicitly assume that no
larger population exist, or that the neurons of the sample are isolated
from their environment.

The \me\ application at the population level gives in fact, by
construction, a distribution with lower Shannon relative entropy than the
one given by the application at the sample level: both satisfy the same
constraints, but only the latter is obtained by direct maximization of the
entropy. The lower entropy of the distribution obtained from application at
the population level is clearly visible from its higher sharpness in
\fig~\ref{fig:diff_maxent_pop_sample}. Such distribution therefore
excludes, probabilistically speaking, more sample states $\yx$ than the
distribution obtained at the sample level does. A lower entropy signals the
presence of additional assumptions or constraints. In this case the
assumption is that the sample is part of a larger population, and our
ignorance about the sample should be representative of our ignorance about
a larger population. This assumption affects our predictions.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-18T14:02:33+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

% \DeclareMathSymbol{\varNu}{\mathalpha}{egreeki}{"4E}

@@

 \DeclareMathSymbol{\varUpsilon}{\mathalpha}{egreeki}{"07}

@@

\newcommand*{\yxx}{x}%subpop state
\newcommand*{\yx}{\bm{\yxx}}%subpop state
\newcommand*{\yxs}{\sav{\yx}}%subpop av state
\newcommand*{\yX}{\bm{X}}%pop state
\newcommand*{\yy}{\bm{y}}%pop state
\newcommand*{\yXf}{\av{\yX}}%pop av state
\newcommand*{\yxxs}{\sav{\yx\yx}}%subpop av state
\newcommand*{\yXXf}{\av{\yX\yX}}%subpop av state
\newcommand*{\yr}{\bm{r}}%subpop value
\newcommand*{\ys}{\bm{s}}%subpop value
\newcommand*{\yrs}{\sav{\yr}}%subpop av value
\newcommand*{\yR}{\bm{R}}%pop value
\newcommand*{\yRf}{\av{\yR}}%pop av value
%conditional assumptions
\newcommand*{\yH}{\varIota}
\newcommand*{\yD}{\varDelta}
\newcommand*{\yHa}{\varIota'}
\newcommand*{\yHb}{\varIota''}
\newcommand*{\yHc}{\varKappa}
\newcommand*{\yHd}{\varUpsilon}
%Lagr multipliers
\newcommand*{\yf}{\bm{f}}
\newcommand*{\yc}{\bm{c}}
\newcommand*{\yL}{\bm{\varLambda}}
\newcommand*{\yl}{\bm{\lambda}}
\newcommand*{\yk}{\kappa}
\newcommand*{\yK}{\varKappa}
%entropies
\newcommand*{\ysh}{H}
\newcommand*{\ybu}{H_{\text{B}}}

@@

A perfunctory intuitive reasoning seems insufficient for clarifying this
point. Let's express it via the probability calculus. Suppose we do not
know whether the sample is really part of a larger population; we do not
know if $N=n$ or how large is $N$ otherwise. Call this state of ignorance
$\yHd$. From the point of view of the probability calculus, this ignorance
about $N$ is expressed by assigning a probability distribution
$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$.
Maintaining our assumption of symmetric ignorance, probability assignments
that do not assume a specific value of $N$ are then obtained by
multiplication of all $N$-dependent probabilities by $\p(N \| \yHd)$ and
subsequent marginalization over $N$ -- technically speaking, $N$ is a
\emph{nuisance parameter} that we eliminate. The probability obtained from
\me\ at the population level, \eqn~\eqref{eq:app_maxent_pop_marg} then
generalizes to
\begin{equation}
  \label{eq:app_maxent_pop_marg_unknown_N}
  \p(\yxs = \yrs \| \yHd)  = \sum_N \biggl\{\yK_N\, 
  \sum_{N\yRf=0}^N \ypp_N(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[{\yL_N}\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
  \,\p(N \| \yHd).
\end{equation}
This is a formidable expression. But the answer to our question, whether
the usual \me\ at the sample level~\eqref{eq:app_maxent_sample} does not
assume anything about a larger population, translates into the precise
mathematical question: are the
distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
\eqref{eq:app_maxent_sample} equal, for some choice of $\p(N \| \yHd)$? We
leave this mathematical problem for future work.

\iffalse
The former application must therefore implicitly assume that no
larger population exist, or that the neurons of the sample are isolated
from their environment.

The \me\ application at the population level gives in fact, by
construction, a distribution with lower Shannon relative entropy than the
one given by the application at the sample level: both satisfy the same
constraints, but only the latter is obtained by direct maximization of the
entropy. The lower entropy of the distribution obtained from application at
the population level is clearly visible from its higher sharpness in
\fig~\ref{fig:diff_maxent_pop_sample}. Such distribution therefore
excludes, probabilistically speaking, more sample states $\yx$ than the
distribution obtained at the sample level does. A lower entropy signals the
presence of additional assumptions or constraints. In this case the
assumption is that the sample is part of a larger population, and our
ignorance about the sample should be representative of our ignorance about
a larger population. This assumption affects our predictions.
\fi
